---
title: "Data analysis for root trait publications"
author: "Lee Ming Yang"
date: '2022-11-09'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This R markdown file extracts out the information from "Frontiers_final (LMY 04042023).xlsx" which contains all details of the publications on roots globally outside the Neotropics. All analysis of the data loaded is also contained within this R.markdown file.

Loads the data 

```{r - Libraries and loading data, warning=FALSE}

library(circlize)
library(readxl)
library(dplyr)
library(stringr)
library(tidyverse)
library(ggplot2)
library(UpSetR)
library(knitr) 
library(kableExtra) 
library(DT)
library(tm)
library(topicmodels)
library(reshape2)
library(pals)
library(SnowballC)
library(lda)
library(ldatuning)
library(flextable)
library(SentimentAnalysis)
library(syuzhet)
library(plotrix)
library(tidyr)
library(bibliometrix)
library(bibliometrixData)
library(kgc)
library(topicmodels)
library(topicdoc)
library(doParallel)
library(LDAvis)
library(servr)
library(Rfast)
library(maps)
library(viridis)
library(viridisLite)
library(tidyr)
library(ggpubr)
library(stringi)

# Loads the raw data
data_pubs <- read_xlsx ("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx", trim_ws = TRUE, skip = 1, na = "NA", col_names = TRUE)

# Narrows down the data to what we need
data_pubs <- data_pubs [,-which(names(data_pubs) %in% c("Source.Title","DOI","Database"))] # Removes journal name, DOI and WoS/Scopus

# Preliminary test if data is clean
# sum(is.na(data_pubs$Article.Title)) # Should be 0
# sum(is.na(data_pubs$Publication.Year)) 
# sum(is.na(data_pubs$Abstract)) 
# sum(is.na(data_pubs$Keywords)) 
# sum(is.na(data_pubs$Country)) 
# sum(is.na(data_pubs$`Type of study`)) 
# sum(is.na(data_pubs$`Sp/Comm`))
# sum(is.na(data_pubs$`CR/FR/both`))
# sum(is.na(data_pubs$Traits)) # all should be 0, which means fully filled

# Rename columns 
colnames (data_pubs)[c(11,14,16:19)] <- c("Site.no","Study_type","Factor","Sp_comm","root_type","FR_class")

```


## Test for trait coverage across publications.

```{r - Trait category analysis}

# First check to see which traits are there (alot, but to check if the abbreviations are consistent)
traits_list <- data_pubs$Traits

# To check if there is any traits that have been entered wrongly
# traits_test <- unlist(strsplit(traits_list, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
# traits_uniq <- unique(traits_test)
# sort(traits_uniq)
# All traits entered correctly
# rm(traits_test,traits_uniq)

# Creates sets of traits for testing of binary categories
biomass_set <- c("BGB","R:S ratio","RMF","FRB","CRB")
morph_set <- c("SRL","SRSA","RTD","RD","RLA","root_length","root_area")
MYC_set <- c("MYC","MYC_status")
phys_set <- c("nitrogenase","exudates","P uptake","phosphatase","root metabolism","N uptake","root_respiration","K uptake")
arch_set <- c("depth","RLD","RBI","RMD","fractal","18O")
dynamics_set <- c("NPP_fineroots","decomposition","FRP","mortality","turnover","necromass","NPP_roots","dynamics (appearance/disappearance","growth","CRP","N:P mineralisation","15N")
chemistry_set <- c("N content","P content","C content","K content","Ca content","Mg content","Mn content","Na content","C:N ratio","N:P ratio")
water_set <- c("hydro_cond","vessel_D","water_uptake")
others_set <- c("starch","NSC","lignin","tannin")

# Empty matrix of yes/no
traits_logic <- matrix(data = 0,nrow = length(traits_list),ncol = 9)
colnames(traits_logic) <- c("Morphology","Mycorrhizal","Physiology","Architecture","Biomass","Dynamics","Chemistry","Water","Others")

# For each publication entry, test if the traits are within one of these categories
# If yes = 1; if no = 0
for (ii in c(1:length(traits_list))){
  
  # Gets the first entry's traits
  traits_hold <- traits_list[ii]
  traits_hold<- unlist(strsplit(traits_hold, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE)) # This delists each trait into individual character strings
  
  # Runs the logic test
  if (sum(traits_hold %in% morph_set) != 0){traits_logic[ii,1] <- 1}
  if (sum(traits_hold %in% MYC_set) != 0){traits_logic[ii,2] <- 1}
  if (sum(traits_hold %in% phys_set) != 0){traits_logic[ii,3] <- 1}
  if (sum(traits_hold %in% arch_set) != 0){traits_logic[ii,4] <- 1}
  if (sum(traits_hold %in% biomass_set) != 0){traits_logic[ii,5] <- 1}
  if (sum(traits_hold %in% dynamics_set) != 0){traits_logic[ii,6] <- 1}
  if (sum(traits_hold %in% chemistry_set) != 0){traits_logic[ii,7] <- 1}
  if (sum(traits_hold %in% water_set) != 0){traits_logic[ii,8] <- 1}
  if (sum(traits_hold %in% others_set) != 0){traits_logic[ii,9] <- 1}

}

# Check if each row has a sum > 0
# rowSums(traits_logic) # All good!
traits_logic <- data.frame(traits_logic)

# GRoot study needs manual entry (global database paper)
x <- which(grepl("GRooT",data_pubs$Article.Title))
traits_logic[x,] <- c(1,1,1,1,1,1,1,0,1) # No water traits in GRooT

# Creates upset plot
upset(traits_logic,nsets = 9,order.by = "freq",sets.x.label = "Publications per trait category",mainbar.y.label = "Trait Intersections",line.size = 1.5,text.scale = c(1.2,1.1,1,1,1.5,0.9), mb.ratio = c(0.5,0.5))

# Counts no. of publications that studied multiple traits
trait_multiple <- rowSums(traits_logic)
sum(trait_multiple > 1) # 174 studies
# Counts how many publications that studied multiple traits AND did not study biomass
trait_nobiomass <- filter(traits_logic,Biomass == 0)
sum(rowSums(trait_nobiomass) > 1) # 29 studies

```

## Test for associated topics of publications.

Preparation of topic modelling process - optimising model parameters.

```{r - Optimising model parameters}

# set options
options(stringsAsFactors = F)         # no automatic data transformation
options("scipen" = 100, "digits" = 4) # suppress math annotation

# Gets abstracts
abs_raw <- data_pubs$Abstract

# Copying from: https://towardsdatascience.com/a-light-introduction-to-text-analysis-in-r-ea291a9865a8
# Corpus
corpus <- SimpleCorpus(VectorSource(abs_raw))

# Cleaning process
dfCorpus <- tm_map(corpus, content_transformer(tolower))
# Transform to lowercase
dfCorpus <- tm_map(dfCorpus, removeNumbers) # Remove numbers
dfCorpus <- tm_map(dfCorpus, removePunctuation, preserve_intra_word_dashes = TRUE) # Remove punctuation
dfCorpus <- tm_map(dfCorpus, removeWords, stopwords("english")) # Remove stop words
dfCorpus <- tm_map(dfCorpus, stemDocument) # Stems words
# Loads and removes extra junk words
junk_words <- unlist(read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",sheet = "Junk Words",col_names = FALSE))
dfCorpus <- tm_map(dfCorpus,removeWords,junk_words)
dfCorpus <- tm_map(dfCorpus, stripWhitespace) # Remove whitespace

# compute document term matrix with terms >= minimumFrequency
minimumFrequency <- 10
DTM <- DocumentTermMatrix(dfCorpus, control = list(bounds = list(global = c(minimumFrequency, Inf))))
# have a look at the number of documents and terms in the matrix
dim(DTM)

# due to vocabulary pruning, we have empty rows in our DTM
# LDA does not like this. So we remove those docs from the
# DTM and the metadata
sel_idx <- slam::row_sums(DTM) > 0
DTM <- DTM[sel_idx, ]
textdata <- abs_raw[sel_idx]

# How many topics? This isn't an easy question to answer, but we can select based on a few criteria (e.g. coherence, exclusivitiy, AIC of model)
# We tried several approaches to estimate k
# Function for parallel processing
calc_AIC <- function(k){
  print(k)
  library(topicmodels)
  # Conducts the LDA on that specified k
  topicModel <- LDA(DTM, k, method="Gibbs", control=list(iter = 1000, verbose = 25, alpha = 0.05))
  # Calculates log likelihood and then saves
  AIC_save <- AIC(topicModel)

}

# Preparation for parallel processing
no_cores <- detectCores() - 1  
cl <- makeCluster(no_cores)  
registerDoParallel(cl)  

# Conducts the loop to test AIC per k value
result_AIC <- foreach(i=2:50) %dopar% calc_AIC(i) # Parallel processing is so MUCH FASTER
# Preparation for graphing
AIC_k <- unlist (result_AIC)
AIC_k <- data.frame(k = c(2:50),AIC_k = AIC_k) # Lowest AIC derived was 18 in our run, but results will vary depending on seed. Generally, expect lowest_k to be somewhere between 17-24.

# Plotting of AIC curve
ggplot (data = AIC_k, aes(x = k)) +
  geom_path(aes(y = AIC_k)) + 
  geom_point(aes(y = AIC_k)) +
  xlab("No. of topics k") + 
  ylab("AIC")

coh_mean <- NA
ex_mean <- NA

# Runs a loop for values of k to judge topic coherence and exclusivity (higher the better)
for (ii in c(2:50)){
  # Set k = no.of topics 
  k <- ii
  # Compute the LDA model, inference via 500 iterations of Gibbs sampling
  topicModel <- LDA(DTM, k, method="Gibbs", control=list(iter = 500, verbose = 25, seed = seed.no, alpha = 0.05))
  # Topic diagnostics
  coherence_score <- mean(topic_coherence(topicModel,DTM,top_n_tokens = 15))
  exclusion_score <- mean(topic_exclusivity(topicModel,top_n_tokens = 15))

  # Saves the means
  coh_mean <- rbind(coh_mean,mean(coherence_score))
  ex_mean <- rbind(ex_mean,mean(exclusion_score))
}

coh_mean <- coh_mean[-1,]
ex_mean <- ex_mean[-1,]

# Plots graph of coherence
df_means <- data.frame(k = c(2:50),coh_mean = coh_mean,ex_mean = ex_mean)

# No discernible trend in topic coherence or exclusivity. 
ggplot(data = df_means, aes (x = k)) +
  geom_path (aes(y = coh_mean)) +
  geom_path (aes(y = ex_mean))

```


Only run this portion after determining optimal value of k, the number of topics. This was determined to be k = 18 in our run, but you may encounter some variation depending on the seed.
The general number of topics suggested by AIC is anywhere between 16-24. We can construct our topic model and link that to traits coverage here.

```{r - Topic modelling}

# Set k = 18, lowest AIC value
k <- 18
seed.no <- 123456789 # For reproducibility
topicModel <- LDA(DTM, k, method="Gibbs", control=list(iter = 1000, verbose = 25, seed = seed.no, alpha = 0.05))

# have a look a some of the results (posterior distributions)
topic_terms <- terms(topicModel, 15)
# Saves into .csv
# write.csv(topic_terms,"K-topic words.csv")

# Names for the 18 topics to be entered manually
# Have a look a some of the results (posterior distributions)
tmResult <- modeltools::posterior(topicModel)
# for every document we have a probaility distribution of its contained topics
theta <- tmResult$topics
# see alpha from previous model
attr(topicModel, "alpha")
attr(topicModel, "beta")# Alpha is generally recommended to be 50/K. We want to reduce alpha here to make the distribution of topics more peaky within a document, since it is unlikely that a single study will cover topics so broadly

# With theta, can obtain the proportion of each abstract occupied by each topic. We assume each abstract to focus on max 2-3 topics, so alpha is set low to get a "peaky" distribution. Then, we select the highest 2-3 proportion topics in each abstract to "group" said publication to 1-3 topics.
# This grouping can be further explored using an upset diagram - similar to the traits in Sec 1.
theta <- data.frame(theta)
colnames(theta) <- sprintf("T%i",c(1:k)) #Just give generic col names for now

theta <- tmResult$topics
# Produces a loop that specifies two criteria for topics: 
# 1) Needs to be >15%
# 2) If multiple topics all >15%, needs to be top 3
for (ii in c(1:nrow(theta))){
  
  # Probabilities in each document
  theta_doc <- theta[ii,]
  # Which topics got 15% probabilities
  theta_15 <- which(theta_doc > 0.15)
  
  # Check how many topics were selected
  # If 3 or less, can proceed to set dummy values
  if (length(theta_15) < 4) {
    theta[ii,theta_15] <- 1 # Set those to 1, rest of the variables can be set to 0 later
  } else { #If 4 or more
    top_prob <-  head(sort(theta_doc, decreasing=TRUE),3) # Arranges and select highest probabilities
    top_ind <- which(theta_doc %in% top_prob) # Gets their indices
    theta[ii,top_ind] <- 1 # Use indices to set topics to 1
    
  }
}

# Then replace all the non-1s into zeroes
theta[theta != 1] <- 0
rowSums(theta) # No. of topics each study looks at (3 or less for all; some have 4 due to close ties)




# Reloads traits_logic and theta (if you want to skip the computational steps above)
theta <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",trim_ws = TRUE, na = "NA", col_names = TRUE,sheet = "theta")
traits_logic <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",trim_ws = TRUE, na = "NA", col_names = TRUE, sheet = "traits")
# Some publications still have 4 topics because of a tie in probabilities. Document 174,261,303,321 have 4 topics

# Explores which of these topics are studied in conjunction
# Needs title of trait categories
# Full column names
colnames(theta) <- c("Global variation",
                     "Experimental responses",
                     "Functional traits",
                     "Litter decomposition",
                     "Root profile (depth)",
                     "Soil respiration",
                     "Allometry",
                     "Ecosystem productivity",
                     "Forests",
                     "Water limitation",
                     "Plant nutrients",
                     "Seedling inoculation",
                     "Light availability",
                     "Species composition",
                     "Mycorrhizal colonisation",
                     "Savannas",
                     "Root dynamics",
                     "Nutrient availability")
theta <- as.data.frame(theta)                     

# Compute an adjacency matrix based on the presence/absence of trait categories and topics 
adj_matrix <- as.matrix(t(theta[,])) %*% as.matrix(traits_logic[,])
adj.matrix <- as.data.frame(adj_matrix)

# Chord diagram
# Gets percentage of each topic
proportion_topic <- (colSums(theta)/352)*100
proportion_topic
topic_label <- sprintf("%.2f%%",proportion_topic)
# Gets percentage of each trait
proportion_trait <- (colSums(traits_logic)/nrow(traits_logic))*100
proportion_trait
trait_label <- sprintf("%.2f%%",proportion_trait)
# Combine percentage labels to loop through each sector
labels_all <- c(topic_label,trait_label)
# Change names of rows
row.names(adj_matrix) <- sprintf("T%i",c(1:ncol(theta)))

grid.col <- c(T1 = "#757C18",
              T2 = "#757C18",
              T3 = "#EAC300",
              T4 = "#A9180D",
              T5 = "#EAC300",
              T6 = "#A9180D",
              T7 = "#234F1E",
              T8 = "#234F1E",
              T9 = "#2EC06D",
              T10 = "#016064",
              T11 = "#EAC300",
              T12 = "#0AE1FE",
              T13 = "#016064",
              T14 = "#231709",
              T15 = "#0AE1FE",
              T16 = "#2EC06D",
              T17 = "#EAC300",
              T18 = "#016064",
              Morphology = "#ced0da",
              Mycorrhizal = "#ced0da",
              Physiology = "#ced0da",
              Architecture = "#ced0da",
              Biomass = "#ced0da",
              Dynamics = "#ced0da",
              Chemistry = "#ced0da",
              Water = "#ced0da",
              Others = "#ced0da")

# Following this arrangement means the need to rearrange labels
index <- c(7,8,3,5,11,17,10,13,18,9,16,12,15,4,6,14,1,2, # Topics
           19,20,21,22,23,24,25,26,27) # Traits
labels_all <- labels_all[index]
# Colour some labels white for easy visibility
col_labels <- c("white","white", # Allometry and productivity
                "black","black","black","black", # Traits
                "white","white","white", # Resource limitation
                "black","black", # Ecosystems
                "black","black", # Mycorrhizal
                "white","white", # Soils
                "white", # Species composition
                "white","white", # Global change
                "black","black","black","black","black",
                "black","black","black","black")

# Plots track
# Set parameters for chord tracks
circos.par(start.degree = 270)
chordDiagram(adj_matrix, 
             annotationTrack = "grid", 
             preAllocateTracks = 1, 
             grid.col = grid.col, # Colour of sectors
             big.gap = 17, # Controls gap between rows and columns
             small.gap = 1.5, # Controls gaps between sectors
             annotationTrackHeight = c(0.2,0.01),
             order = c("T7","T8",
                       "T3","T5","T11","T17",
                       "T10","T13","T18",
                       "T9","T16",
                       "T12","T15",
                       "T4","T6",
                       "T14",
                       "T1","T2",
                       "Morphology","Mycorrhizal","Physiology",
                       "Architecture","Biomass","Dynamics","Chemistry",
                       "Water","Others")) # Controls thickness of borders

# Runs loop to enter each sector and place label
for(ii in c(1:length(labels_all))) {
  # x and y coordinates of each sector
  xlim = get.cell.meta.data("xlim", sector.index = get.all.sector.index()[ii], track.index = 1)
  ylim = get.cell.meta.data("ylim", sector.index = get.all.sector.index()[ii], track.index = 1)
  # Enters text
  circos.text(mean(xlim), mean(ylim)-1.4, # Need to cut down by 1.4 to keep percentage labels in cell
              labels_all[ii], 
              sector.index = get.all.sector.index()[ii], # Retrieves the sector names, but is not shown in the diagram (edited manually)
              track.index = 1, 
              facing = "clockwise", 
              niceFacing = TRUE, #
              col = col_labels[ii], 
              adj = c(0,0.5),
              cex = 0.6) # Size
}

circos.clear()

```

Codes to count number of publications of specific criteria for in-text numbers.

```{r - Counting topics}

print((colSums(theta))) # Proportion of each topic

# Counting filter by topic category
topic_biomass <- theta %>% filter(Allometry != 0|`Ecosystem productivity`!= 0) #83 studies
topic_function <- theta %>% filter(`Functional traits` != 0|
                                     `Root profile (depth)` != 0|
                                     `Root dynamics` != 0|
                                     `Plant nutrients` != 0) #162 studies
topic_MYC <- theta %>% filter(`Mycorrhizal colonisation` != 0|
                                `Seedling inoculation` != 0) #86 studies
topic_resources <- theta %>% filter(`Water limitation` != 0|
                                      `Nutrient availability` != 0|
                                      `Light availability` != 0) # 130 studies
topic_soil <- theta %>% filter(`Soil respiration` != 0|
                                 `Litter decomposition` != 0) #64 studies
topic_ecosystems <- theta %>% filter(`Forests` != 0|
                                       `Savannas` != 0) #89 studies
topic_global <- theta %>% filter(`Global variation` != 0|
                                   `Experimental responses` != 0) #78 studies
topic_sp <- theta %>% filter(`Species composition` != 0) #23 studies

```

## Ancillary information of publications.

How is trait information represented? We ask:
- Is trait information usually at the species or community-level?
- What environmental variables are of interest in experimental studies?
- What environments are usually studied in in-situ sites?
- What is the criteria used for fine/coarse roots used?

```{r - Breakdown into different study types}

# Classify studies based on type
study_type <- data.frame(type = data_pubs$Study_type,env = data_pubs$Env,var = data_pubs$Factor)
study_type <- study_type %>% mutate(Meta = str_detect(type,"Meta"),
                                    insitu = str_detect(type,c("Field_obs|Field_manipulation|Field_grown|Exp_plots_I")),
                                    exsitu = str_detect(type,c("Glasshouse|Greenhouse|Shadehouse|Potted|Nursery|Exp_plots_X")))
# Count
sum(study_type$Meta) # 45 studies
sum(study_type$insitu) # 234 studies
sum(study_type$exsitu) # 78; 4 both

# Count how many studies are species/community
sp_comm <- as.data.frame(data_pubs$Sp_comm)
colnames(sp_comm) <- "Sp_comm"
sp_comm <-  mutate(sp_comm,Sp = grepl("Species",Sp_comm,fixed = TRUE), Comm = grepl("Community",Sp_comm,fixed = TRUE))

# Merge both dataframes with the raw data
breakdown.df <- cbind(data_pubs$Article.Title,sp_comm,study_type,data_pubs$root_type,data_pubs$FR_class)
colnames(breakdown.df)[c(11:12)] <- c("CR/FR","Criteria")
# Filter by study type
insitu_breakdown <- breakdown.df %>% filter(insitu == TRUE)
exsitu_breakdown <- breakdown.df %>% filter(exsitu == TRUE)
meta_breakdown <- breakdown.df %>% filter(Meta == TRUE)

```

```{r - Sp/Comm}

# For in-situ studies, is data species/community level?
sum(insitu_breakdown$Sp) #91 studies
sum(insitu_breakdown$Comm) # 145 studies; 3 overlap
# Pie chart - 88 species; 142 communities; 3 both
slices <- c(88,142,3) 
par(bg=NA)
pie(slices,labels = NA,col = c("#3DED97","#74B72E","#354A21"),cex = 2)

# For ex-situ studies, all studies should be species-based 
sum(exsitu_breakdown$Sp) # 78 studies; obvious since the experiment is manipulated manually
# Pie chart - 78 for ex-situ studies (all species based)
slices <- c(78)
par(bg=NA)
pie(slices,labels = NA,col = c("#BDD7EE"),cex = 2)

# For meta studies, is data species/community level?
sum(meta_breakdown$Sp) # 6 studies
sum(meta_breakdown$Comm) # 40 studies, 1 overlap

# Pie chart - 5,39,1 for meta analyses type studies
slices <- c(5,39,1)
par(bg=NA)
pie(slices,labels = NA,col = c("#C5C6D0","#59515E","#232023"),cex = 2)

```

```{r - Insitu and meta environments}

# For in-situ studies, which environment?
insitu_env<- unlist(strsplit(insitu_breakdown$env, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(insitu_env)
insitu_env.df <- data.frame(Env = c("Forest","Grassland","Heath","Savanna","Disturbance","Agroforest","Mangrove","Others"),Count = c(163,12,11,22,24,8,16,14))
insitu_env.df$Env <- factor(insitu_env.df$Env,levels = c("Forest","Disturbance","Savanna","Mangrove","Grassland","Heath","Agroforest","Others"),ordered = T)

ggplot(data=insitu_env.df) +
  geom_col(aes(x = Env, y = Count)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Environment") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18))

# For meta-analysis, which env?
meta_env<- unlist(strsplit(meta_breakdown$env, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(meta_env)
meta_env.df <- data.frame(Env = c("Forest","Grassland","Savanna","Mangrove","Others"),Count = c(29,9,5,2,13))
meta_env.df$Env <- factor(meta_env.df$Env, levels = c("Forest","Grassland","Savanna","Mangrove","Others"),ordered = T)
                          
ggplot(data=meta_env.df) +
  geom_col(aes(x = Env, y = Count)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Environment") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18))
                  
```

```{r - Exsitu variables}

# For ex-situ studies, which variables?
exsitu_var <- unlist(strsplit(exsitu_breakdown$var, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(exsitu_var)

# Creates long format for ggplotting of variables tested
# Disturbance includes fire, flooding and pruning
# Nutrients include nutrients, N and P
# Others include all other variables
exsitu_var.df <- data.frame(Variable = c("Light","Water","Disturbance","Nutrients","Temperature","CO2","Mycorrhizal","Others"),
                            Count = c(21,12,8,21,1,3,15,24)) 
exsitu_var.df$Variable <- factor(exsitu_var.df$Variable,levels =
c("Nutrients","Light","Mycorrhizal","Water","Disturbance","CO2","Temperature","Others"),ordered = T)                                  
exp1 <- expression("CO"[2])
ggplot(data=exsitu_var.df) +
  geom_col(aes(x = Variable, y = Count)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  scale_x_discrete(labels = c("Nutrients","Light","Mycorrhizal","Water","Disturbance",exp1,"Temperature","Others")) +
  xlab("Variable manipulated") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18))

```

```{r - Criteria for distinguishing FR/CR in in-situ studies}

# Breakdown of in-situ studies (whether they study CR/FR) and how they distinguished
# Unfortunately, most studies were unclear if they distinguished between the 2
# First, filters out the studies that clearly employed such a criteria to distinguish
insitu_disting <- insitu_breakdown %>% filter(Criteria != "Unknown" & !is.na(Criteria)) # 116 studies
# This also means 233-116 = 117 studies did not distinguish CR/FR clearly
# Among the 116 studies, did they study CR/FR/both?
insitu_CRFR <- unlist(strsplit(insitu_disting$`CR/FR`, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(insitu_CRFR) #39 both; 2 CR only; 75 FR

# Pie chart - 39 both,2 CR,75 FR,117 unknown
slices <- c(39,75,2,117)
par(bg=NA)
pie(slices,labels = NA,col = c("#dc731c","#7eb7e8","#257675","#808080"),cex = 2)

# Among those that did, what criteria did they use?
insitu_crit <- unlist(strsplit(insitu_disting$Criteria, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
criteria.df <- as.data.frame(table(insitu_crit))
criteria.df$insitu_crit <- factor(criteria.df$insitu_crit,levels = c("<10mm","<5mm","<3mm","<2mm","<1mm","Diameter class","Order"),ordered = T)

ggplot(data=criteria.df) +
  geom_col(aes(x = insitu_crit, y = Freq)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Fine-root criteria") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  scale_x_discrete(labels = c("<10mm","<5mm","<3mm","<2mm","<1mm","Diameter class","Order"))

```

```{r - Criteria for distinguishing FR/CR in ex-situ studies}

# Breakdown of ex-situ studies (whether they study CR/FR) and how they distinguished
# Unfortunately, most studies were unclear if they distinguished between the 2
# First, filters out the studies that clearly employed such a criteria to distinguish
exsitu_disting <- exsitu_breakdown %>% filter(Criteria != "Unknown" & !is.na(Criteria)) # 9 studies
# This also means 78-9 = 69 studies did not distinguish CR/FR clearly
# Among the 9 studies, did they study CR/FR/both?
exsitu_CRFR <- unlist(strsplit(exsitu_disting$`CR/FR`, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(exsitu_CRFR) #4 both; 5 FR

# Pie chart - 4 both, 5 FR, 69 unknown
slices <- c(4,5,69)
par(bg=NA)
pie(slices,labels = NA,col = c("#dc731c","#7eb7e8","#808080"),cex = 2)

# Among those that did, what criteria did they use?
exsitu_crit <- unlist(strsplit(exsitu_disting$Criteria, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
criteria.df <- as.data.frame(table(exsitu_crit))
criteria.df$exsitu_crit <- factor(criteria.df$exsitu_crit,levels = c("<2mm","Order"),ordered = T)

ggplot(data=criteria.df) +
  geom_col(aes(x = exsitu_crit, y = Freq)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Fine-root criteria") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18))

```

```{r - Criteria for distinguishing FR/CR in meta studies}

# Breakdown of ex-situ studies (whether they study CR/FR) and how they distinguished
# Unfortunately, most studies were unclear if they distinguished between the 2
# First, filters out the studies that clearly employed such a criteria to distinguish
meta_disting <- meta_breakdown %>% filter(Criteria != "Unknown" & !is.na(Criteria)) # 18 studies
# This also means 45-18 = 27 studies did not distinguish CR/FR clearly
# Among the 18 studies, did they study CR/FR/both?
meta_CRFR <- unlist(strsplit(meta_disting$`CR/FR`, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
table(meta_CRFR) #6 both; 1 CR; 11 FR

# Pie chart - 6 both, 1 CR, 11 FR, 27 unknown
slices <- c(6,11,1,27)
par(bg=NA)
pie(slices,labels = NA,col = c("#dc731c","#7eb7e8","#257675","#808080"),cex = 2)

# Among those that did, what criteria did they use?
meta_crit <- unlist(strsplit(meta_disting$Criteria, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
criteria.df <- as.data.frame(table(meta_crit))
criteria.df$meta_crit <- factor(criteria.df$meta_crit,levels = c("<5mm","<2mm","Diameter_class","functional_class","Order","All"),ordered = T)

ggplot(data=criteria.df) +
  geom_col(aes(x = meta_crit, y = Freq)) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1)) +
  xlab("Fine-root criteria") +
  ylab("No. of studies") +
  theme(rect = element_rect(fill = "transparent"),
        text = element_text(size = 18)) +
  scale_x_discrete(labels = c("<5mm","<2mm","Diameter class","Functional class","Order","Multiple")) +
  scale_y_continuous(breaks = seq(from = 0, to = 16, by = 2))

```



## Geographical coverage of traits and topics.

Plots the world map of sites. 

```{r - Plots world map}

# Plotting of studies on a world map only makes sense for in-situ studies (aka field studies)
# Classify studies based on type
study_type <- data.frame(type = data_pubs$Study_type,env = data_pubs$Env,var = data_pubs$Factor)
study_type <- study_type %>% mutate(Meta = str_detect(type,"Meta"),
                                    insitu = str_detect(type,c("Field_obs|Field_manipulation|Field_grown|Exp_plots_I")),
                                    exsitu = str_detect(type,c("Glasshouse|Greenhouse|Shadehouse|Potted|Nursery|Exp_plots_X")))
# Count
sum(study_type$Meta) # 45 studies
sum(study_type$insitu) # 233 studies
sum(study_type$exsitu) # 78 studies; 4 both 
# write.csv(study_type,"C:/Users/chaos/Desktop/Network Analysis (Chapter 1)/study_type.csv") # To save

# Traits binary
traits_logic <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",trim_ws = TRUE, na = "NA", col_names = TRUE, sheet = "traits")
# Topics binary
topics_logic <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",trim_ws = TRUE, na = "NA", col_names = TRUE,sheet = "theta")
# Study type
study_type <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx",trim_ws = TRUE, na = "NA", col_names = TRUE,sheet = "study_type")
# Pulls out individual sites data
multi_sites <- read_xlsx("C:/Users/chaoslifes/Desktop/Frontiers_Word_Templates/Frontiers_final (LMY 04042023) - Cleaned data.xlsx", sheet = "Multi_locations", trim_ws = TRUE, na = "NA", col_names = TRUE) # This last sheet is used as a pivot table for studies with multiple field sites

# Creates a new dataframe for coordinates, location, traits and topics
geog_df <- data_pubs[,c("Article.Title","Publication.Year","DOI_name","Lat","Long","Site.no","Country")]
geog_df <- cbind(geog_df,study_type,traits_logic,topics_logic)
# Filters out insitu studies (b/c it makes sense to map out those studies on a map)
geog_map <- geog_df %>% filter(insitu == "TRUE")
  
# Preps all the datapoints to plot first
# First, gets all the studies with multiple sites
geog_multi <- geog_map %>% dplyr::filter(Site.no > 1)  #Multiple sites
geog_single <- geog_map %>% filter(Site.no == 1) # Single sites

# Retrieves coordinates of all the sites from the multisites list
colnames(multi_sites)[1] <- "DOI_name"
coord_multi <- multi_sites %>% filter(DOI_name %in% geog_multi$DOI_name)
coord_multi <- coord_multi[,c(1:3)]
# Then combines with the geog_single coordinates
geog_single <- geog_single[,c(3,4,5)] # Single sites only
coord_total <- rbind(geog_single,coord_multi) # This is a dataframe to plot sites ONLY 

# Preps the dataframe that decides the colours of each country
# Colours will denote the no. of studies within each country
countries_raw <- geog_map$Country
# Replace Borneo(I) with Indonesia; replace Borneo(M) with Malaysia
countries_raw <- gsub("Borneo(M)","Malaysia",countries_raw,fixed = TRUE)
countries_raw <- gsub("Borneo(I)","Indonesia",countries_raw,fixed = TRUE)
# Delist to count
countries_delist <- unlist(strsplit(countries_raw, "; ", fixed = FALSE, perl = FALSE, useBytes = FALSE))
countries_count <- count(countries_delist)
colnames(countries_count)[1] <- "region"

# Generic worldmap
world <- map_data("world")
# Merge with countries count
world_count <- left_join(world,countries_count,by = "region") # Many countries will probably have no studies; they should take a freq of 0
world_count$freq[is.na(world_count$freq)] <- 0
# Splits the dataframe for countries with actual numerical values for colouring
world_nonzero <- world_count %>% filter(freq > 0)

ggplot(world_count, aes(long, lat, group = group))+
  geom_polygon(fill = "grey",colour = "white",alpha = 0.5) +
  geom_polygon(data = world_nonzero, aes(x = long, y = lat, group = group, fill = freq),colour = "white",inherit.aes = FALSE) +
  theme_void() +
  xlim(-30,max(world_count$long)) +
  ylim(-45,max(world_count$lat)) +
  geom_point(data = coord_total,aes(x = Long,y = Lat),inherit.aes = FALSE, color = "red", show.legend = TRUE, size = 1.5) +
  geom_rect(aes(xmin = -25,
                xmax = 170,
                ymin = -35,
                ymax = 35),colour = "black", fill = NA) +
  labs(fill="No. of studies") +
  scale_fill_gradientn(limits = c(0,50),
                       breaks = seq(0,50,10),
                       colours = c("#AFEAE4","#051937")) +
  theme(legend.title=element_text(margin=margin(0,0,7,0)),
        legend.text=element_text(size = 7))

```

Geographical trait coverage.

```{r - Geographical trait coverage}

# Obtains statistics of the in-situ studies for each geographical region
# First area - Africa
geog_africa <- geog_map %>% filter(Country %in% c("Ghana","Nigeria","Ivory Coast","Tanzania","Cameroon","Democratic Republic of the Congo","Uganda","Madagascar","Kenya","Ethiopia","Republic of Congo","Central Africa Republic","Burkina Faso","Rwanda","Zimbabwe","Senegal")) #45 studies
africa_traits <- geog_africa[,c(14:22)]
africa_topics <- geog_africa[,c(23:40)]
colSums(africa_traits)
(colSums(africa_traits)/nrow(africa_traits))*100
colSums(africa_topics)
(colSums(africa_topics)/nrow(africa_topics))*100

# 2nd area - India and Nepal
geog_india <- geog_map %>% filter(Country %in% c("India","Nepal")) # 51 observations
india_traits <- geog_india[,c(14:22)]
india_topics <- geog_india[,c(23:40)]
colSums(india_traits)
(colSums(india_traits)/nrow(india_traits))*100
colSums(india_topics)
(colSums(india_topics)/nrow(india_topics))*100

# 3rd area - China
geog_china <- geog_map %>% filter(Country %in% c("China")) # 27 observations
china_traits <- geog_china[,c(14:22)]
china_topics <- geog_china[,c(23:40)]
colSums(china_traits)
(colSums(china_traits)/nrow(china_traits))*100
colSums(china_topics)
(colSums(china_topics)/nrow(china_topics))*100

# 4th area - SEA
geog_SEA <- geog_map %>% filter(Country %in% c("Borneo(M)","Indonesia; Thailand","Indonesia","Thailand","Philippines","Vietnam","Singapore","Malaysia","Brunei","Borneo(I)","Cambodia")) # 87 observations
SEA_traits <- geog_SEA[,c(14:22)]
SEA_topics <- geog_SEA[,c(23:40)]
colSums(SEA_traits)
(colSums(SEA_traits)/nrow(SEA_traits))*100
colSums(SEA_topics)
(colSums(SEA_topics)/nrow(SEA_topics))*100

# 5th area - Australia and Oceania
geog_aus <- geog_map %>% filter(Country %in% c("Australia","Micronesia","Papua New Guinea","New Caledonia")) # 23 observations
aus_traits <- geog_aus[,c(14:22)]
aus_topics <- geog_aus[,c(23:40)]
colSums(aus_traits)
(colSums(aus_traits)/nrow(aus_traits))*100
colSums(aus_topics)
(colSums(aus_topics)/nrow(aus_traits))*100

# Traits dataframe
traits_geog <- rbind (colSums(africa_traits),colSums(india_traits), colSums(china_traits),colSums(SEA_traits),colSums(aus_traits))
geog_regions <- c("Africa","India","China","SEA","Aus")  
# Recorders locations
traits_geog <- cbind(geog_regions,traits_geog)
# Converts to long format
traits_geog <- as.data.frame(traits_geog)
traits_long <- pivot_longer(data = traits_geog,
                            cols = -geog_regions,
                            names_to = "traits",
                            values_to = "studies_no.")
traits_long$studies_no. <- as.numeric(traits_long$studies_no.)

# Presents the above information in a balloon plot
# Too many traits/topics categories; thus use balloon plot for easier visualisation as compared to a bar chart
# For traits information first
traits_long$traits <- str_replace_all(traits_long$traits,"Water_traits","Water")
traits_long$geog_regions <- str_replace_all(traits_long$geog_regions,"Aus","Oceania")
traits_long$geog_regions <- factor(traits_long$geog_regions,levels = c("Africa","India","China","SEA","Oceania"),ordered = T)

# Balloon plot
my_cols <- c("#F0F921FF","#E16462FF","#B12A90FF","#6A00A8FF","#0D0887FF")
ggballoonplot(data = traits_long,fill = "value") +
  scale_fill_gradientn(colors = my_cols,
                       name = "No. of studies") +
  labs(size="") 

```

Geographical topics coverage.

```{r - Geographical topics coverage}

# Topics dataframe
topics_geog <- rbind (colSums(africa_topics),colSums(india_topics), colSums(china_topics),colSums(SEA_topics),colSums(aus_topics))
colnames(topics_geog) <- c("Global variation",
                     "Experimental responses",
                     "Functional traits",
                     "Litter decomposition",
                     "Root profile (depth)",
                     "Soil respiration",
                     "Allometry",
                     "Ecosystem productivity",
                     "Forests",
                     "Water limitation",
                     "Plant nutrients",
                     "Seedling inoculation",
                     "Light availability",
                     "Species composition",
                     "Mycorrhizal colonisation",
                     "Savannas",
                     "Root dynamics",
                     "Nutrient availability")
geog_regions <- c("Africa","India","China","SEA","Aus")
topics_geog <- cbind(geog_regions,topics_geog)

# Converts to long format for plotting
topics_geog <- as.data.frame(topics_geog)
topics_long <- pivot_longer(data = topics_geog,
                            cols = -geog_regions,
                            names_to = "topics",
                            values_to = "studies_no.")


topics_long$geog_regions <- str_replace_all(topics_long$geog_regions,"Aus","Oceania")
topics_long$geog_regions <- factor(topics_long$geog_regions,levels = c("Africa","India","China","SEA","Oceania"),ordered = T) # Recorders locations

topics_long$topics <- factor(topics_long$topics,levels = c("Experimental responses","Global variation","Species composition","Soil respiration","Litter decomposition","Mycorrhizal colonisation","Seedling inoculation","Savannas","Forests","Nutrient availability","Light availability","Water limitation","Root dynamics","Plant nutrients","Root profile","Functional traits","Ecosystem productivity","Allometry"),ordered = T) # Reorders topics

# Creates the plot
topics_long$studies_no. <- as.numeric(topics_long$studies_no.)
ggballoonplot(data = topics_long, fill = "value") +
  scale_fill_gradientn(colors = my_cols,
                       name = "No. of studies") +
  labs(size="") 


```


## Temporal coverage of traits and topics. 

Temporal breakdown of traits.

```{r - Temporal breakdown of traits}

# Renames topics by X1-18 for easier referencing
colnames(geog_df)[c(23:40)] <- sprintf("X%i",c(1:18))

# Breakdown of studies by years
# old == before 2000
# past == 2001-2009
# new == 2010-2015
# modern == 2016 onwards
pubs_old <- geog_df %>% filter(Publication.Year <= 2000) #76 studies
pubs_past <- geog_df %>% filter(Publication.Year <= 2009 & Publication.Year >= 2001) #81 studies
pubs_new <- geog_df %>% filter(Publication.Year <= 2015 & Publication.Year >= 2010) #71 studies
pubs_modern <- geog_df %>% filter(Publication.Year > 2015) #124 studies

# 2000 and before
old_traitsrank <- apply(pubs_old[,c(14:22)],2,sum)
old_traitsrank
old_topicsrank <- apply(pubs_old[,c(23:40)],2,sum)
old_topicsrank

# 2001-2009
past_traitsrank <- apply(pubs_past[,c(14:22)],2,sum)
past_traitsrank
past_topicsrank <- apply(pubs_past[,c(23:40)],2,sum)
past_topicsrank

# 2010-2015
new_traitsrank <- apply(pubs_new[,c(14:22)],2,sum)
new_traitsrank
new_topicsrank <- apply(pubs_new[,c(23:40)],2,sum)
new_topicsrank

# 2015-2021
modern_traitsrank <- apply(pubs_modern[,c(14:22)],2,sum)
modern_traitsrank
modern_topicsrank <- apply(pubs_modern[,c(23:40)],2,sum)
modern_topicsrank

# Creates vector for time periods
vect_periods <- c("Before 2000","2001-2009","2010-2015","2016-present")

# Binds all sums of traits and topics by periods
temporal_traitdf <- rbind(old_traitsrank,past_traitsrank,new_traitsrank,modern_traitsrank)
temporal_traitdf <- as.data.frame(cbind(vect_periods,temporal_traitdf))
row.names(temporal_traitdf) <- c(1:4)
colnames(temporal_traitdf)[9] <- "Water"

# Transforms into long df for ggplot
temporal_traitdf <- pivot_longer(data = temporal_traitdf,
                                 cols = -vect_periods,
                                 names_to = "Traits",
                                 values_to = "Count")
colnames(temporal_traitdf)[1] <- "Time_period"
temporal_traitdf$Count <- as.numeric(temporal_traitdf$Count)
# Ascribes order to periods
temporal_traitdf$Time_period <- factor(temporal_traitdf$Time_period, levels = c("Before 2000","2001-2009","2010-2015","2016-present"))

# Count for total no. of studies by period
temporal_totaldf <- data.frame(Time_period = c("Before 2000","2001-2009","2010-2015","2016-present"), Count = c(76,81,71,124)) 

# Barplot + line graph
# Grey bars in the background indicate no. of studies in total
# Lines denote no. of studies by each trait category
# Note that many studies have more than 1 trait category
t1 <- ggplot(data = temporal_traitdf) +
  geom_line(aes(x = Time_period, y = Count, group = Traits, color = Traits)) +
  geom_point(aes(x = Time_period, y = Count, group = Traits, color = Traits), size = 2.5, shape = 18) +
  geom_col(data = temporal_totaldf,aes(x = Time_period, y = Count),color = "grey",alpha = 0.15) +
  xlab("Time period") +
  ylab("No. of studies") +
  theme_classic() +
  scale_y_continuous(limits = c(0,130), expand = c(0,0)) +
  scale_color_manual(values = c('#00429d','#81436b','#a36667',
                                '#a8a695','#00ffff','#cad085',
                                '#c6ad97','#9f9b73','#469400'),
                        name = "Trait categories") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.position = "none")

```

Temporal breakdown of topics.

```{r - Temporal breakdown of topics}

# Prepares dataframes for topics
# Presents information by topic categories because 18 topics is too much
# For publications before 2000
old_topics_sum <- pubs_old %>% mutate(Biomass_prod = case_when(X7 != 0 ~ TRUE,
                                                                 X8 != 0 ~ TRUE,
                                                                 TRUE ~ FALSE),
                                        Func_traits = case_when(X3 != 0 ~ TRUE,
                                                                X5 != 0 ~ TRUE,
                                                                X11 != 0 ~ TRUE,
                                                                X17 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Resource_ava = case_when(X10 != 0 ~ TRUE,
                                                                X13 != 0 ~ TRUE,
                                                                X18 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Eco_dyn = case_when(X9 != 0 ~ TRUE,
                                                            X16 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Sp_comp = case_when(X14 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Mycorrhizal_test = case_when(X12 != 0 ~ TRUE,
                                                            X15 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Soil_test = case_when(X4 != 0 ~ TRUE,
                                                            X6 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Env_change = case_when(X1 != 0 ~ TRUE,
                                                            X2 != 0 ~ TRUE,
                                                            TRUE ~ FALSE))
old_topics_sum <- old_topics_sum %>% dplyr::select(Biomass_prod:Env_change)                               

# For publications 2001-2009
past_topics_sum <- pubs_past %>% mutate(Biomass_prod = case_when(X7 != 0 ~ TRUE,
                                                                 X8 != 0 ~ TRUE,
                                                                 TRUE ~ FALSE),
                                        Func_traits = case_when(X3 != 0 ~ TRUE,
                                                                X5 != 0 ~ TRUE,
                                                                X11 != 0 ~ TRUE,
                                                                X17 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Resource_ava = case_when(X10 != 0 ~ TRUE,
                                                                X13 != 0 ~ TRUE,
                                                                X18 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Eco_dyn = case_when(X9 != 0 ~ TRUE,
                                                            X16 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Sp_comp = case_when(X14 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Mycorrhizal_test = case_when(X12 != 0 ~ TRUE,
                                                            X15 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Soil_test = case_when(X4 != 0 ~ TRUE,
                                                            X6 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Env_change = case_when(X1 != 0 ~ TRUE,
                                                            X2 != 0 ~ TRUE,
                                                            TRUE ~ FALSE))
past_topics_sum <- past_topics_sum %>% dplyr::select(Biomass_prod:Env_change)                                                                
# For publications 2010-2015
new_topics_sum <- pubs_new %>% mutate(Biomass_prod = case_when(X7 != 0 ~ TRUE,
                                                                 X8 != 0 ~ TRUE,
                                                                 TRUE ~ FALSE),
                                        Func_traits = case_when(X3 != 0 ~ TRUE,
                                                                X5 != 0 ~ TRUE,
                                                                X11 != 0 ~ TRUE,
                                                                X17 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Resource_ava = case_when(X10 != 0 ~ TRUE,
                                                                X13 != 0 ~ TRUE,
                                                                X18 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Eco_dyn = case_when(X9 != 0 ~ TRUE,
                                                            X16 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Sp_comp = case_when(X14 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Mycorrhizal_test = case_when(X12 != 0 ~ TRUE,
                                                            X15 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Soil_test = case_when(X4 != 0 ~ TRUE,
                                                            X6 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Env_change = case_when(X1 != 0 ~ TRUE,
                                                            X2 != 0 ~ TRUE,
                                                            TRUE ~ FALSE))
new_topics_sum <- new_topics_sum %>% dplyr::select(Biomass_prod:Env_change)   

# For publications 2016 onwards
modern_topics_sum <- pubs_modern %>% mutate(Biomass_prod = case_when(X7 != 0 ~ TRUE,
                                                                 X8 != 0 ~ TRUE,
                                                                 TRUE ~ FALSE),
                                        Func_traits = case_when(X3 != 0 ~ TRUE,
                                                                X5 != 0 ~ TRUE,
                                                                X11 != 0 ~ TRUE,
                                                                X17 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Resource_ava = case_when(X10 != 0 ~ TRUE,
                                                                X13 != 0 ~ TRUE,
                                                                X18 != 0 ~ TRUE,
                                                                TRUE ~ FALSE),
                                        Eco_dyn = case_when(X9 != 0 ~ TRUE,
                                                            X16 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Sp_comp = case_when(X14 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Mycorrhizal_test = case_when(X12 != 0 ~ TRUE,
                                                            X15 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Soil_test = case_when(X4 != 0 ~ TRUE,
                                                            X6 != 0 ~ TRUE,
                                                            TRUE ~ FALSE),
                                        Env_change = case_when(X1 != 0 ~ TRUE,
                                                            X2 != 0 ~ TRUE,
                                                            TRUE ~ FALSE))
modern_topics_sum <- modern_topics_sum %>% dplyr::select(Biomass_prod:Env_change) 

# Now, we can safely sum within each category and bind them
temporal_topicdf <- as.data.frame(rbind(colSums(old_topics_sum),colSums(past_topics_sum),colSums(new_topics_sum),colSums(modern_topics_sum)))

# Adds periods
temporal_topicdf <- as.data.frame(cbind(vect_periods,temporal_topicdf))
row.names(temporal_topicdf) <- c(1:4)

# Transforms into long df for ggplot
temporal_topicdf <- pivot_longer(data = temporal_topicdf,
                                 cols = -vect_periods,
                                 names_to = "Topics",
                                 values_to = "Count")
colnames(temporal_topicdf)[1] <- "Time_period"
temporal_topicdf$Count <- as.numeric(temporal_topicdf$Count)
# Ascribes order to periods
temporal_topicdf$Time_period <- factor(temporal_topicdf$Time_period, levels = c("Before 2000","2001-2009","2010-2015","2016-present"))
# Ascribes order to topic categories 
temporal_topicdf$Topics <- factor(temporal_topicdf$Topics ,
                                  levels = c("Biomass_prod","Env_change","Eco_dyn","Mycorrhizal_test","Func_traits","Soil_test","Resource_ava","Sp_comp"),ordered = T)

# Barplot + line graph
# Grey bars in the background indicate no. of studies in total
# Lines denote no. of studies by each trait category
# Note that many studies have more than 1 trait category
t2 <- ggplot(data = temporal_topicdf) +
  geom_line(aes(x = Time_period, y = Count, group = Topics, color = Topics)) +
  geom_point(aes(x = Time_period, y = Count, group = Topics, color = Topics), size = 2.5, shape = 18) +
  geom_col(data = temporal_totaldf,aes(x = Time_period, y = Count),color = "grey",alpha = 0.15) +
  xlab("Time period") +
  ylab("No. of studies") +
  theme_classic() +
  scale_y_continuous(limits = c(0,130), expand = c(0,0)) +
  scale_color_manual(values = c("#234F1E","#757C18","#2EC06D","#0AE1FE","#EAC300","#A9180D","#016064","#231709"),
                   name = "Topic categories",
                  labels = c("Biomass and productivity","Global ecosystem variation","Ecosystem-specific dynamics","Mycorrhizal","Functional traits","Soil processes","Resource availability","Biodiversity")) +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.position = "none") # Remove this line for legend

# For counting purposes
# old_topics <- subset(pubs_old, select = X1:X18)
# colnames(old_topics) <- c("Global variation",
#                      "Experimental responses",
#                      "Functional traits",
#                      "Litter decomposition",
#                      "Root profile",
#                      "Soil respiration",
#                      "Allometry",
#                      "Ecosystem productivity",
#                      "Forests",
#                      "Water limitation",
#                      "Plant nutrients",
#                      "Seedling inoculation",
#                      "Light availability",
#                      "Species composition",
#                      "Mycorrhizal colonisation",
#                      "Savannas",
#                      "Root dynamics",
#                      "Nutrient availability")
# colSums(old_topics)
# past_topics <- subset(pubs_past, select = X1:X18)
# colnames(past_topics) <- c("Global variation",
#                      "Experimental responses",
#                      "Functional traits",
#                      "Litter decomposition",
#                      "Root profile",
#                      "Soil respiration",
#                      "Allometry",
#                      "Ecosystem productivity",
#                      "Forests",
#                      "Water limitation",
#                      "Plant nutrients",
#                      "Seedling inoculation",
#                      "Light availability",
#                      "Species composition",
#                      "Mycorrhizal colonisation",
#                      "Savannas",
#                      "Root dynamics",
#                      "Nutrient availability")
# colSums(past_topics)
# new_topics <- subset(pubs_new, select = X1:X18)
# colnames(new_topics) <- c("Global variation",
#                      "Experimental responses",
#                      "Functional traits",
#                      "Litter decomposition",
#                      "Root profile",
#                      "Soil respiration",
#                      "Allometry",
#                      "Ecosystem productivity",
#                      "Forests",
#                      "Water limitation",
#                      "Plant nutrients",
#                      "Seedling inoculation",
#                      "Light availability",
#                      "Species composition",
#                      "Mycorrhizal colonisation",
#                      "Savannas",
#                      "Root dynamics",
#                      "Nutrient availability")
# colSums(new_topics)
# modern_topics <- subset(pubs_modern, select = X1:X18)
# colnames(modern_topics) <- c("Global variation",
#                      "Experimental responses",
#                      "Functional traits",
#                      "Litter decomposition",
#                      "Root profile",
#                      "Soil respiration",
#                      "Allometry",
#                      "Ecosystem productivity",
#                      "Forests",
#                      "Water limitation",
#                      "Plant nutrients",
#                      "Seedling inoculation",
#                      "Light availability",
#                      "Species composition",
#                      "Mycorrhizal colonisation",
#                      "Savannas",
#                      "Root dynamics",
#                      "Nutrient availability")
# colSums(modern_topics)

# Merge into single image for publication
ggarrange(t1 + rremove("xlab") + rremove("x.text"), 
          t2, ncol = 1, nrow = 2,
              common.legend = FALSE,
              align = "hv", 
              font.label = list(size = 8, color = "black", face = "bold",family = NULL))

```


